Computing in the Field of Rationals

### 2012-06-21 rational_ext.tex
_

\begin{proposition}[Strong Normalizing]
\label{prop:termination-pi} 
{{Pi}} computation always terminate.

@@

We present the categorical structure with minimum commentry. More
details may be found in excellent references such as Barr and Wells
[CITE] and Selinger[CITE].

@@

  \begin{proof}
    To establish that category is monoidal one must show isomorphisms
    \begin{enumerate}
    \item {{b1 + (b2 + b3) <-> (b1 +b2) + b3}} given by {{assoc+}}
    \item {{0 + b <-> b}} given by {{zeroe}}
    \item {{b + 0 <-> b}} given by {{swap+ (;) zeroe}}
    \item We need to check that {{+}} is a bifunctor. 
      \begin{enumerate}
      \item {{id_{b1<->b1}(+)id_{b2<->b2} = id_{b1+b2<->b1+b2} }}
      \item {{(f(+)g)(;)(j(+)k) = (f (;) j) (+) (g (;) k)}}
      \end{enumerate}
    \item We need to check the naturality of {{assoc+}}, {{zeroe}} and
      {{swap+(;)zeroe}}.
      \begin{enumerate}
      \item {{assoc+ (;) ((f (+)g) (+) h) = (f (+)(g (+) h)) (;) assoc+}}
      \item {{zeroe (;) f = (id (+) f) (;) zeroe}}
      \item {{(swap+ (;) zeroe) (;) f = (f (+) id) (;) (swap+ (;) zeroe)}}
      \end{enumerate}
    \item Satisfy certain coherence conditions which are usually
      called the ``pentagon'' and ``triangle'' axioms.
    \end{enumerate}

    The last three points require checking equality of combinators by
    writing out their derivation trees as we did in the case of
    associativity of sequential composition. To show symmetry, we need
    a braiding operation {{b1+b2 <-> b2+b1}} which is given by
    {{swap+}}.

    \begin{enumerate}
    \item The braiding must satisfy two ``hexagon axioms''. 
    \item The braiding is self inverse {{swap+_{b1+b2} = (swap+_{b2+b1})^{dagger} }}.
    \end{enumerate}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Traces and the Int-Construction.}

>>>>

\begin{proposition}[Strong Normalizing]
\label{prop:termination-pi} 
{{Pi}} computations always terminate.

@@

We present the categorical structure with minimum commentry. More
details may be found in excellent references such as Barr and Wells
[CITE] and Selinger \cite{springerlink:10.1007/978-3-642-12821-94}.

@@

  \begin{proof}
    To establish that category is monoidal one must show isomorphisms
    \begin{enumerate}
    \item {{b1 + (b2 + b3) <-> (b1 +b2) + b3}} given by {{assoc+}}
    \item {{0 + b <-> b}} given by {{zeroe}}
    \item {{b + 0 <-> b}} given by {{swap+ (;) zeroe}}
    \item We need to check that {{+}} is a bifunctor. 
      \begin{enumerate}
      \item {{id_{b1<->b1}(+)id_{b2<->b2} = id_{b1+b2<->b1+b2} }}
      \item {{(f(+)g)(;)(j(+)k) = (f (;) j) (+) (g (;) k)}}
      \end{enumerate}
    \item We need to check the naturality of {{assoc+}}, {{zeroe}} and
      {{swap+(;)zeroe}}.
      \begin{enumerate}
      \item {{assoc+ (;) ((f (+)g) (+) h) = (f (+)(g (+) h)) (;) assoc+}}
      \item {{zeroe (;) f = (id (+) f) (;) zeroe}}
      \item {{(swap+ (;) zeroe) (;) f = (f (+) id) (;) (swap+ (;) zeroe)}}
      \end{enumerate}
    \item Satisfy certain coherence conditions which are usually
      called the ``pentagon'' and ``triangle'' axioms (see Sec 3.1
      \cite{springerlink:10.1007/978-3-642-12821-94})
    \end{enumerate}

    The last three points require checking equality of combinators by
    writing out their derivation trees as we did in the case of
    associativity of sequential composition. To show symmetry, we need
    a braiding operation {{b1+b2 <-> b2+b1}} which is given by
    {{swap+}} (see Sec. 3.3 and 3.5
    \cite{springerlink:10.1007/978-3-642-12821-94}).

    \begin{enumerate}
    \item The braiding must satisfy two ``hexagon axioms''.
    \item The braiding is self inverse
      {{swap+_{b1+b2}=(swap+_{b2+b1})^{dagger} }}.
    \end{enumerate}

@@

\item To be pedantic, what we have shown is that the ``term model'' of
  {{Pi}} that follows from the extensional operational equality of
  combinators has the requisite categorical structure. A consequence
  is that the ``wiring diagrams'' of {{Pi}} correspond closely with ``string
  diagrams'' developed for categories.

  To establish the later rigorously, we will need to show when it is
  valid to slide one wire over the other and that equivalent diagrams
  for syntactically different combinators such as
  {{(f(+)g)(;)(j(+)k)}} and {{(f(;)j)(+)(g(;)k)}} do respect
  operational equivalence.  We don't formalize the graphical notation
  in this work. Joyal et. al's work on ``planar isotopy'' and
  Selinger's survey \cite{springerlink:10.1007/978-3-642-12821-94}
  show how this has been addressed before in the categorical
  setting. In the absense of any prior knowledge of category theory
  however, our wiring diagrams may be read as the ``flow of types'' in
  a combinator-circuit.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Traces and the Int-Construction.}
\label{sec:int}
